### 2022-02-12 paper/output/document/thesis_3_implementation.tex
making progress on CHAPTER 3

%
\begin{isabellebody}%
\setisabellecontext{paper{\isadigit{4}}{\isadigit{1}}}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Novel Formalization of the Categorical Imperative%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this section, I present a custom formalization of the categorical imperative, as inspired by 
the goals from the previous chapter.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Logical Background%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
The previous attempts to model the categorical imperative in Chapter 2 partially failed due to 
an inability to fully represent the complexity of a maxim. Specifically, they treated actions as a single, 
monolithic unit of evaluation, whereas most Kantians consider the unit of evaluation for the FUL to the more
complex notion of a maxim. In this section, I will present some logical background necessarily to fully 
capture the spirit of a maxim. I will begin by borrowing some machinery to handle ``subjects" who perform 
actions from Chapter 2.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{typedecl}\isamarkupfalse%
\ s\ %
\isamarkupcmt{s is the type for a ``subject," i.e. the subject of a sentence. In this interpretation, 
a subject is merely defined as ``that which can act." It does not include any other properties, such as 
rationality or dignity. As I will show, for the purposes of defining the universalizability test, this 
``thin" representation of a subject suffices.%
}\isanewline
\isanewline
\isacommand{type{\isacharunderscore}synonym}\isamarkupfalse%
\ os\ {\isacharequal}\ {\isachardoublequoteopen}{\isacharparenleft}s\ {\isasymRightarrow}\ t{\isacharparenright}{\isachardoublequoteclose}\ %
\isamarkupcmt{Recall that an open sentence maps a subject to a term to model the 
substitution operator.%
}\isanewline
\isanewline
\isacommand{type{\isacharunderscore}synonym}\isamarkupfalse%
\ maxim\ {\isacharequal}\ {\isachardoublequoteopen}{\isacharparenleft}t\ {\isacharasterisk}\ os\ {\isacharasterisk}\ t{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
The central unit of evaluation for the universalizability test is a ``maxim," which Kant defines 
in a footnote in \emph{Groundwork} as ``the subjective principle of willing," or the principle that 
the agent acts on \cite[16]{groundwork}. Modern Kantians differ in their interpretations of this definition. The naive view 
is that a maxim is an act, but Korsgaard adopts the more sophisticated view that a maxim is composed
of an act and the agent's purpose for acting \cite{actingforareason}. She also compares a maxim 
to Aristotle's logos, which includes these components and information about the circumstances and methods 
of the act. O'Neill concludes that Kant's examples imply that a maxim must also include circumstances \cite{actingonprinciple}, and 
Kitcher \cite{whatisamaxim} uses textual evidence from the Groundwork to argue for the inclusion of a maxim's purpose 
or motivation. In order to formalize the notion of a maxim, I must adopt a specific definition and 
defend my choice.

I define a maxim as a circumstance, act, goal tuple (C, A, G), read 
as ``In circumstances C, act A for goal G." Isabelle's strict typing rules mean that the choice of the 
type of each member of this tuple is significant. A circumstance is represented as a set of worlds 
$t$ where that circumstance holds. A goal is also a term because it can be true or false at a world if it 
is realized or not. An act is an open sentence because an act itself is not the kind of thing that can 
be true or false (as in, an act is not truth-apt), but the combination of a subject performing an act 
can be true or false at a world depending on whether or not the act is indeed performed by that subject. 
For example, ``running" is not truth-apt, but ``Sara runs" is truth-apt.

My definition of a maxim is inspired by O'Neill's work on maxims. I will defend my representation
below and consider an additional component that Kitcher argues for.

$\emph{O'Niell's Original Schematic and The Role of Practical Judgement}$

O'Neill \cite[37]{actingonprinciple} presents what Kitcher \cite{whatisamaxim}  calls the widely accepted 
view that a maxim is a circumstance, act, goal tuple. A maxim 
is an action-guiding rule and thus naturally includes an act and the circumstances under which 
it should be performed, which are often referred to as ``morally relevant circumstances."

She also includes a purpose, end, or goal in the maxim because Kant includes this in many of his 
example maxims and because Kant argues that human activity, because it is guided by a rational will, 
is inherently purposive \cite[4:428]{groundwork}. A rational will does not act randomly (else it would not be rational), 
but instead in the pursuit of ends which it deems valuable. This inclusion is also essential for the version of the universalizability test 
that I will implement, explained in Section ??.

O'Neill's inclusion of circumstances is potentially controversial because it leaves open the question of what qualifies as a 
relevant circumstance for a particular maxim. This is gives rise to ``the tailoring objection" \cite[217]{whatisamaxim} \footnote{Kitcher
cites \cite{kantsethicalthought}  as offering an example of a false positive due to this objection.}, 
under which maxims are arbitrarily specified to pass the FUL. For example, the maxim ``When my name is Lavanya Singh,
I will lie to get some easy money," is universalizable, but is clearly a false positive. One solution to 
this problem is to argue that the circumstance ``When my name is Lavanya Singh" is not morally relevant 
to the act and goal. This solution requires some discussion of what qualifies as a relevant circumstance.

O'Neill seems to acknowledge the difficulty of determining relevant circumstances when she concedes that a maxim cannot include all 
of the infinitely many circumstances in which the agent may perform the action\cite[4:428]{actingonprinciple}. She argues that this is 
an artifact of the fact that maxims are rules of practical reason, the kind of reason that helps us decide what to do 
and how to do it \cite{bok}. Like any practical rule, 
maxims require the exercise of practical judgement to determine in which circumstances they should be applied. 
This judgement, applied in both choosing when to exercise the maxim and in the formulation of the maxim 
itself, is what determines the ``morally relevant circumstances."

The upshot for computational ethics is that the computer cannot perform all ethical activity alone. 
Human judgement and the exercise of practical reason are essential to both formulate maxims and 
determine when the actual conditions of life coincide with the circumstances in which the maxim is relevant. 
Choosing when to exercise a maxim is less relevant to my project because analyzing a formal representation of the FUL requires 
making the circumstances in a given scenario precise, but will be important for applications of 
computational ethics to guiding AI agents. The difficulty in formulating a maxim, on the other hand, demonstrates 
the important fact that ethics, as presented here, is not a solely computational activity. A
human being must create a representation for the dilemma they wish to test, effectively translating 
a complex, real situation into a flat logical structure. This parallels the challenge that programmers 
face when translating the complexity of reality to a programming langauge or computational representation. Not only will some of the situation's complexity
inevitably be lost, the outcome of the universalizability test will depend on how the human formulates the maxim
and whether or not this formulation does indeed include morally relevant circumstances. If the human puts 
garbage into the test, the test will return garbage out.

While this may appear to be a weakness of my system, I believe that it actually
allows my system to retain some of the human complexity that many philosophers agree cannot be automated away.\footnote{Powers presents 
the determination of morally relevant circumstances as an obstacle to the automation of Kantian ethics \cite{powers}.}
Ethics is a fundamentally human activity. Kant argues that the categorical imperative is a statement 
about the properties of rational wills. In fact, Korsgaard argues that morality derives its authority over us, 
or normativity, only because is it a property of a rational will, and we, as human beings, are rational wills.
If ethics is meant to guide human behavior, the role of the computer becomes clear as not a replacement for our will,
but instead as a tool to help guide our wills and reason more efficiently 
and more effectively. Just as calculators don't render mathematicians obsolete, computational ethics
does not render human judgement or philosophy obsolete. Chapter 4 Section ?? will be devoted to a more complete discussion 
of this issue.

$\emph{Exclusion of Motive}$

Kitcher begins with O'Niell's circumstance, act, goal view and expands it to include the motive 
behind performing the maxim \cite{whatisamaxim}. This additional component is read 
as ``In circumstance C, I will do A in order to G because of M," where M may be ``duty" or ``self-love."
Kitcher argues that the inclusion of motive is necessary for the fullest, most general form of a maxim
in order to capture Kant's idea that an action derives its moral worth from being done for the sake of duty itself.
Under this view, the FUL would obligate maxims of the form 
``In circumstance C, I will do A in order to G because I can will that I and everyone else simultaneously
will do A in order to G in circumstance C." In other words, if Kant is correct in arguing that moral 
actions must be done from the motive of duty, the affirmative result of the FUL becomes 
the motive for a moral action.

While Kitcher's conception of a maxim captures Kant's idea of acting for duty's own sake, I will not implement it 
because it is not necessary for putting maxims through the FUL. Indeed, Kitcher acknowledges that 
O'Neill's formulation suffices for the universalizability test, but is not the general notion of a maxim.
In order to pass the maxim through the FUL, it suffices to know the circumstance, act, and goal. The FUL
derives the motive that Kitcher bundles into the maxim, so automating the FUL does not require 
including a motive. The ``input" to the FUL is the circumstance, act, goal tuple. My project takes 
this input and returns the motivation that the dutiful, moral agent would adopt. Additionally, doing
justice to the rich notion of motive requires modelling the operation of practical reason itself, 
which is outside the scope of this project. My work focuses on the universalizability test, but future work that 
models the process of practical reason may use my implementation of the FUL as a ``library." Combined 
with a logic of practical reason, an implementation of the FUL can move from evaluating a maxim to 
evaluating an agent's behavior, since that's when ``acting from duty" starts to matter.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ will\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}maxim\ {\isasymRightarrow}\ s{\isasymRightarrow}\ \ t{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}W\ {\isacharunderscore}\ {\isacharunderscore}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \isakeyword{where}\ {\isachardoublequoteopen}will\ {\isasymequiv}\ {\isasymlambda}{\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}\ s{\isachardot}\ {\isacharparenleft}c\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}a\ s{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
Korsgaard claims that ``to will an end, rather than just
wishing for it or wanting it, is to set yourself to be its cause" \cite[38]{sources}. To will a maxim
is to set yourself to be the cause of its goal by taking the means 
specified in the maxim in the relevant circumstances. This coheres with 
Kitcher's and Korsgaard's understanding of a maxim as a principle or rule to live by.

At worlds
where the circumstances do not hold, a maxim is vacuously willed. If you decide to act on the rule ``I will 
do X in these cirumstances", then you are vacuously obeying it when the circumstances don't hold.

The above discussion implies that willing a maxim is particular to the agent, justifying my choice to 
require that a particular subject will a maxim. O'Neill argues for this interpretation when she distinguishes 
between the evaluation of a principle, which is generic, and a maxim, which she views as ``individuated only 
by referring to a person"\cite[13]{actingonprinciple}. I adopt the spirit of this interpretation but modify it slightly 
by representing the general maxim as a principle that anyone could adopt, and the act of willing the maxim 
as a person-particular instantiation of the maxim.

I additionally represent a subject as willing a maxim because I use the word `will' as a verb, to mean committing oneself to living by
the principle of a maxim. This coheres with the FUL, which tests the act willing 
of a maxim by determining if the maxim could be a universal law that everyone committed to. Formalizing this idea,
the type of a willed maxim is a term, allowing me
to use DDL's obligation operator on the notion of willing a maxim. Concretely, my system will prove 
or disprove statements of the form ``Lavanya is obligated to will the maxim M."

Worlds where the circumstances do not hold are not relevant for determining obligation. Recall that in 
Benzmueller et. al's definition of the obligation operator,  $O \{B|A\}$ is true at all worlds iff ob(B)(A), or 
if the obligation function maps A to obligatory in context B (where the context is a set of worlds) \cite{BFP}. This 
definition implies that worlds outside of B have no bearing on the moral status of A in context B, which 
coheres with intuitions about contextual obligation. Thus, the dyadic obligation operator 
disqualifies worlds where the context does not hold, so the vacuous truth of the will statement in 
these worlds does not matter.

Given that the will abbreviation already excludes worlds where the circumstances fail (by rendering 
the statement vacuously true at them), one may conclude that the dyadic obligation operator is now useless. 
Using the dyadic obligation operator allows me to take advantage of the power of DDL to represent the bearing 
that circumstances have on obligation. DDL has powerful axioms expressing the relationship between circumstances 
and obligation, such as the fact that obligations are monotonically increasing with respect to broader 
circumstances. Using the monadic obligation operator would require me to either operate with an empty 
notion of context or to redefine these axioms. The dyadic obligation operator lets me take advantage of the full 
power of DDL in expressing contrary-to-duty obligations. This is particularly important for Kantian ethics 
and the FUL specifically because many critiques of the FUL rely on attention to circumstances (tailoring 
objection) or lack thereof (ideal theory). This is also an innovation that my custom formalization presents 
over the prior work. By formally including the notion of a circumstance or context, I am able to represent 
these objections that Kantian scholars study. Formalizing Kantian ethics in a dyadic deontic logic 
instead of a monadic deontic logic is a key contribution of this thesis.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ effective\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}maxim{\isasymRightarrow}s{\isasymRightarrow}\ t{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}E\ {\isacharunderscore}\ {\isacharunderscore}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \isakeyword{where}\ {\isachardoublequoteopen}effective\ \ {\isasymequiv}\ {\isasymlambda}{\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}\ s{\isachardot}\ {\isacharparenleft}{\isacharparenleft}will\ {\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}\ s{\isacharparenright}\ \isactrlbold {\isasymequiv}\ g{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isacommand{print{\isacharunderscore}theorems}\isamarkupfalse%
%
\begin{isamarkuptext}%
A maxim is effective for a subject when, if the subject wills it then the goal is achieved, and
when the subject does not act on it, the goal is not achieved$\footnote{Thank you to Jeremy D. Zucker for helping me think through this.}$ \cite{sepcausation}. 
The former direction of the implication 
is intuitive: if the act results in the goal, it was effective in causing the goal. This represents `necessary'
causality.

The latter direction represents `sufficient' causality, or the idea that, counterfactually,
if the maxim were not willed, then the goal is not achieved \cite{lewiscausality}. Note that nothing else changes about this
counterfactual world—the circumstances are identical and we neither added additional theorems nor 
specified the model any further. This represents Lewis's idea of "comparative similarity,"  where 
a counterfactual is true if it holds at the most similar world \cite{lewiscounterfactuals}. In our case, this is just the world 
where everything is the same except the maxim is not acted on.

Combining these ideas, this definition of effective states that a maxim is effective if the 
maxim being acted on by a subject is the necessary and sufficient cause of the goal.\footnote{Should I wave a hand at critiques of counterfactual causality?}

If the circumstances do not hold and the goal is achieved, then the maxim is vacuously effective, since 
it is vacuously willed (as described above). While this scenario is counterintuitive, it is not very 
interesting for my purposes because, when the circumstances do not hold, a maxim is not applicable. It 
doesn't really make sense to evaluate a maxim when it's not supposed to be applied. The maxim ``When on Jupiter,
read a book to one-up your nemesis" is vacuously effective because it can never be disproven.%

@@

@@

%:%file=~/Desktop/cs91r/paper/paper41.thy%:%

@@

%:%46=11%:%
%:%58=13%:%
%:%59=14%:%
%:%60=15%:%
%:%61=16%:%

@@

@@

%:%66=20%:%
%:%67=20%:%
%:%68=21%:%
%:%69=22%:%
%:%70=23%:%
%:%71=23%:%
%:%72=24%:%
%:%73=25%:%
%:%74=25%:%
%:%75=25%:%
%:%76=26%:%
%:%77=26%:%
%:%78=27%:%
%:%79=28%:%
%:%80=28%:%
%:%82=30%:%
%:%83=31%:%
%:%84=32%:%
%:%85=33%:%
%:%86=34%:%
%:%87=35%:%
%:%88=36%:%
%:%89=37%:%
%:%90=38%:%
%:%91=39%:%
%:%92=40%:%
%:%93=41%:%
%:%94=42%:%
%:%95=43%:%
%:%96=44%:%
%:%97=45%:%
%:%98=46%:%
%:%99=47%:%
%:%100=48%:%
%:%101=49%:%
%:%102=50%:%
%:%103=51%:%

>>>>

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{3}}{\isacharunderscore}implementation}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Implementation Details\label{details}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this section, I present the details of my implementation of automated Kantian ethics. I take
a logic-programming approach in which I formalize the FUL in Dyadic Deontic Logic and then implement 
this logic in Isabelle/HOl. I 
also present my testing framework, which demonstrates one way to evaluate how faithful an implementation
of automated ethics is to philosophical literature. This testing framework shows that my system outperforms
two other possible implementations of automated Kantian Ethics.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Formalization and Implementation of the FUL%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Before formalizing the FUL, I must define and implement the relevant logical background. Dyadic
Deontic Logic can express obligation and prohibition, but it cannot represent more complex features of 
moral judgement like actions, subject, maxims, and ends. I augment DDL by adding representations 
of these concepts, drawn from philosophical literature.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Logical Background%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Kantian ethics is centered on practical reason because it is action-guiding; the categorical
imperative is a moral rule that agents can use to decide between potential actions. Thus, before I even
begin to formalize a specific formulation of the categorical imperative, I must define the notions of 
subjects and actions. Concretely, I need to add logical background so that my logic can express sentences 
like, ``x does action.''%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{typedecl}\isamarkupfalse%
\ s\ %
\isamarkupcmt{I declare a new type, $s$ as the type for a ``subject,'' i.e. the subject of a sentence.%
}%
\begin{isamarkuptext}%
The \texttt{typedecl} keyword indicates that I am defining a new atomic type, which is not composed
of pre-existing types but is instead a new kind of thing altogether. I try to minimize the number of 
atomic types, so this will be one of the few new types that I define. To define a subject, it suffices
to declare a new type to represent a subject, as that is all that is needed to apply the Formula of
Universal Law. Notice that I have not defined any properties of this type, such as the idea that a subject
must be rational or human. Throughout my project, I will use bare syntactic units like types and constants 
to define new ideas and will add the minimum necessary definitions, or ``thin'' definitions, to achieve the desired results.
By avoiding complex definitions of a subject, I can avoid murky philosophical debates about the nature
of agency and reduce the potential for controversy or errors.

In this interpretation, the defining feature of a subject is that it can act, a relatively uncontroversial
notion. I represent that below by allowing subjects to substitute into sentences, a property that I will use 
to represent the notion of different people performing certain actions.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{type{\isacharunderscore}synonym}\isamarkupfalse%
\ os\ {\isacharequal}\ {\isachardoublequoteopen}{\isacharparenleft}s\ {\isasymRightarrow}\ t{\isacharparenright}{\isachardoublequoteclose}\ %
\isamarkupcmt{To model the idea of a subject being substituted into an action, I 
define \texttt{type\_synonym} $os$ for an open sentence. An open sentence takes as input a subject and 
returns a complete or ``closed'' DDL formula by, binding the free variable in the sentence to the input. 
For example, ``runs'' is an open sentence that can be instantiated with subject, ``Sara'' to create
the DDL term ``Sara runs,'' which can be true or false at a world. An open sentence itself is not
the kind of thing that can be true or false at a world, so it is not truth-apt, but when an action is 
substituted into an open sentence, the resulting term is truth apt. ``Runs'' is not the kind of thing 
that can be true or false, but ``Sara runs'' is a sentence that can be true or false.%
}%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Maxim \label{whatisamaxim}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Recall from Section \ref{whyful} that I formalize a version of the categorical imperative called
the Formula of Universal, which reads ``act only according to that maxim by which you can at the same 
time will that it should become a universal law'' \cite{groundwork}. In order to faithfully formalize
the FUL, I must make precise the notions of of ``willing a maxim'' and violating the FUL. I draw on 
reliable definitions of willing, maxims, and the FUL from Kantian literature and represent them in DDL.

The central unit of evaluation for Kantian ethics is a ``maxim,'' which Kant defines as ``the subjective 
principle of willing,'' or the principle that the agent understands themselves as acting on \cite{groundwork}. 
Modern Kantians differ in their interpretations of this definition. I adopt O'Neill's view, derived from 
Kant's example maxims, that a maxim includes the act, the circumstances, and the agent's purpose of 
acting or goal \citep{actingonprinciple}.

\begin{definition}[Maxim]
    A maxim is a circumstance, act, goal tuple (C, A, G), read as ``In circumstances C, act A for goal G.''
\end{definition}

I implement this definition in Isabelle by defining the \texttt{type\_synonym} below for the type
of a maxim.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{type{\isacharunderscore}synonym}\isamarkupfalse%
\ maxim\ {\isacharequal}\ {\isachardoublequoteopen}{\isacharparenleft}t\ {\isacharasterisk}\ os\ {\isacharasterisk}\ t{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{A maxim is of type term, open sentence, term tuple, such as ``(When I am broke, will falsely promise
to repay a loan, to get some easy cash)''. The first term represents the circumstance, which
can be true or false at a world. For example, the circumstance ``when I am broke'' is true at the real 
world when my bank account is empty. The second term represents the action, which is an open sentence
because different agents can perform a particular action. For example, the action, ``will falsely promise
to repay a loan'' is an open sentence that can be acted on by a subject. The third term represents 
the goal, which can again be true or false at a world. For example, the goal ``to get some easy cash''
is true at the real world if I have gotten easy cash.%
}%
\begin{isamarkuptext}%
O'Neill \cite[37]{actingonprinciple} argues that maxim 
is an action-guiding rule and thus naturally includes an act and the circumstances under which 
it should be performed, which are often referred to as ``morally relevant circumstances'' \citep[37]{actingonprinciple}. 
She also includes a purpose, end, or goal in the maxim because human activity, is guided by a rational will, 
and is thus inherently purposive \citep[4:428]{groundwork}. A rational will does not act randomly (else it would not be rational), 
but instead in the pursuit of ends which it deems valuable\footnote{Some argue that a maxim should also
include the agent's motive or motivation and I address this concern in Appendix \ref{maximmotive}.}. The inclusion a maxim's end is essential for the version of the FUL
that I will implement, explained in Section Practical Contradiction.

\color{red} SHOULD THIS GO HERE OR IN THE LIMITATIONS SECTION OF THE DISCUSSION
O'Neill's inclusion of circumstances is potentially controversial because it leaves open the question of what qualifies as a 
relevant circumstance for a particular maxim. This is gives rise to ``the tailoring objection" \citep[217]{whatisamaxim} \footnote{Kitcher
cites \citet{kantsethicalthought} as offering an example of a false positive due to this objection.}, 
under which maxims are arbitrarily specified to pass the FUL. For example, the maxim ``When my name is Jane Doe,
I will lie to get some easy money," is universalizable, but is clearly a false positive. One solution to 
this problem is to argue that the circumstance ``When my name is Jane Doe" is not morally relevant 
to the act and goal. This solution requires determining what qualifies as a relevant circumstance.

O'Neill seems to acknowledge the difficulty of determining relevant circumstances when she concedes that a maxim cannot include all 
of the infinitely many circumstances in which the agent may perform an action \citep[4:428]{actingonprinciple}. She argues that this is 
an artifact of the fact that maxims are rules of practical reason, the kind of reason that helps us decide what to do 
and how to do it \citep{bok}. Like any practical rule, 
maxims require the exercise of practical judgement to determine in which circumstances they should be applied. 
This judgement, applied in both choosing when to exercise the maxim and in the formulation of the maxim 
itself, is what determines the morally relevant circumstances.

The difficulty in determining relevant circumstances is a limitation of my system and may require that a 
human being formulate the maxim or that future work develop heuristics to classify circumstances as morally 
relevant. For proponents of the ``human-in-the-loop'' model of AI ethics, in which ethical AI requires that 
humans guide machines, this kind of human involvement may be a feature \citep{loop}. In this model, 
a human being must create a representation for the dilemma they wish to test, translating 
a complex situation into a flat logical structure. This parallels the challenge that programmers 
face when translating the complexity of reality to a programming langauge or computational representation. 
The outcome of the universalizability test will depend on how the human formulates the maxim
and whether or not this formulation does indeed include morally relevant circumstances. If the human puts 
garbage into the test, the test will return garbage out.

Another solution to this problem may be to develop heuristics to classify circumstances as morally 
relevant. For example, one such attempt could define a moral closeness relation between an action, a 
goal, and circumstances. This heuristic would define morally relevant circumstance as those that 
reach a certain closeness threshhold with the action and the goal. Determining morally relevant
circumstances, either using heuristics or human involvement, is a ripe area for future work.
\color{black}

To will a maxim is to adopt it as a principle to live by, or to commit oneself to the action for the 
sake of the end in the relevant circumstances. Korsgaard argues that ``to will an end, rather than 
just wishing for it or wanting it, is to set yourself to be its cause" \cite[38]{sources}. I formalize
this idea in Definition \ref{willing}.

\begin{definition}[Willing]\label{willing}
For maxim $M = (C, A, G)$ and actor $s$,

$$will \, M \, s \equiv \forall w \, (C \longrightarrow A \, (s)) \, w$$

At all worlds $w$, if the circumstances hold at that world, agent $s$ performs act $A$.

\end{definition}

If I will the example maxim above about falsely promising to pay a loan, then whenever I need cash, 
I will falsely promise to repay a loan. I can represent this definition using the following Isabelle
formula.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ will\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}maxim\ {\isasymRightarrow}\ s{\isasymRightarrow}\ \ t{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}W\ {\isacharunderscore}\ {\isacharunderscore}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \isakeyword{where}\ {\isachardoublequoteopen}will\ {\isasymequiv}\ {\isasymlambda}{\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}\ s{\isachardot}\ {\isacharparenleft}c\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}a\ s{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{An agent $s$ wills a maxim iff in the circumstances, $s$ performs the action, or $s$ substituted
into the open sentence $a$ is true. This is an Isabelle \texttt{abbreviation}, which is syntactic
sugar for an Isabelle formula. The type of this formula is $maxim \rightarrow s \rightarrow t$, so it 
takes as input a maxim and a subject and returns the term, ``s wills maxim.''%
}%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Practical Contradiction Interpretation \label{praccon}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In order to actually evaluate the moral status of a maxim, I must precise the idea of failing
the universalizability test or a non-universalizable maxim.  Kantians debate the correct interpretation of 
the Formula of Universal Law because Kant himself appears to interpret the criterion in different ways. 
My project uses Korsgaard's practical contradiction interpretation, broadly accepted as correct within 
the philosophical community \citep{ebelsduggan}.

Recall that the Formula of Universal Law is to “act only in accordance with that maxim through which 
you can at the same time will that it become a universal law” \citep{groundwork}. To determine if a 
maxim can be willed as a universal law, we can use the “universalizability test”, which requires 
imagining a world in which everyone has willed the maxim. If willing the maxim in such a world 
generates a contradiction, then the action is prohibited.

One interpretation of the FUL, the logical contradiction interpretation, prohibits maxims that are 
logically impossible when universalized. Under this view, falsely promising to repay a loan fails the 
universalizability test because, in the universalized world, the practice of promising would die out, 
so making a false promise would be impossible.

One problem with this view is that it cannot handle natural acts. Korsgaard appeals to 
\citet{dietrichson} to construct the example natural act of a mother killing her children that tend to 
cry at night so that she can get some sleep. Universalizing this maxim does not generate a logical 
contradiction, but it is clearly wrong. Because killing is a natural act, it can never be logically 
impossible so the logical contradiction view cannot prohibit it.

As an alternative to the logical contradiction view, Korsgaard endorses the practical contradiction view, 
which prohibits maxims that are self-defeating, or ineffective, when universalized. By willing a maxim, 
an agent commits themselves to the maxim's goal, and thus cannot rationally will that this goal be 
undercut. This interpretation can prohibit natural acts like those of the sleep-deprived mother: in 
willing the end of sleeping, she is implicitly willing that she is alive. If all mothers kill all 
loud children, then she cannot be secure in the possession of her life, because her own mother would 
have killed her as an infant. Her willing this maxim thwarts the end that she sought to secure.

The practical contradiction interpretation offers a satisfying explanation of \emph{why} certain 
maxims are immoral. These maxims involve parasitic behavior on social conditions that the agent seeks 
to benefit from. The false promiser simultaneously wants to abuse the system of promising and benefit 
from it, and is thus making an exception of themselves. The test formalizes the kinds of objections 
that the question ``what if everyone did that?'' seeks to draw out.

Using the practical contradiction interpretation, the FUL states, ``If, when universalized, a maxim is 
not effective, then it is prohibited.'' This requires defining effectiveness and universalization. 
If an agent wills an effective maxim, then the maxim's goal is achieved, and if the agent does 
not will it, then the goal is not achieved.

\begin{definition}[Effective Maxim]
For a maxim $M = (C, A, G)$ and actor $s$,

$$\text{\emph{effective}} \, M \, s \equiv \forall w \, (\text{\emph{will}} \, (C, A, G) \, s \iff G) \, w$$

\end{definition}

I can implement this in Isabelle using another abbreviation.%
\ effective\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}maxim{\isasymRightarrow}s{\isasymRightarrow}\ t{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}E\ {\isacharunderscore}\ {\isacharunderscore}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \isakeyword{where}\ {\isachardoublequoteopen}effective\ \ {\isasymequiv}\ {\isasymlambda}{\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}\ s{\isachardot}\ {\isacharparenleft}{\isacharparenleft}will\ {\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}\ s{\isacharparenright}\ \isactrlbold {\isasymequiv}\ g{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isamarkupcmt{A maxim is effective for a subject if the goal is achieved if and only if the subject wills the maxim.
Once again, I use an abbreviation to conveniently refer to this Isabelle formula.%
The former direction of the implication is intuitive: if the act results in the goal, it was 
effective in causing the goal. This represents sufficient causality. The latter direction represents 
necessary causality, or the idea that, counterfactually, if the maxim were not willed, then the goal 
would not be achieved \citep{lewiscausality}.\footnote{Thank you for Jeremy Zucker for helping me 
think about causality in this way.}  Note that nothing else changes about this
counterfactual world—the circumstances are identical and we neither added additional theorems nor 
specified the model any further. This represents Lewis's idea of "comparative similarity,"  where 
a counterfactual is true if it holds at the most similar world \cite{lewiscounterfactuals}. 
In our case, this is just the world where everything is the same except the maxim is not acted on.
Combining these ideas, this definition of effective states that a maxim is effective if the 
maxim being acted on by a subject is the necessary and sufficient cause of the goal.%

@@

\isacommand{end}\isamarkupfalse%

@@

%:%file=~/Desktop/cs91r/paper/thesis_3_implementation.thy%:%

@@

%:%38=10%:%
%:%39=11%:%
%:%40=12%:%
%:%41=13%:%
%:%50=15%:%

@@

%:%64=19%:%

@@

%:%74=22%:%
%:%86=24%:%
%:%87=25%:%
%:%88=26%:%
%:%89=27%:%
%:%90=28%:%
%:%92=30%:%
%:%93=30%:%
%:%94=30%:%
%:%97=32%:%
%:%98=33%:%
%:%99=34%:%
%:%100=35%:%
%:%101=36%:%
%:%102=37%:%
%:%103=38%:%
%:%104=39%:%
%:%105=40%:%
%:%106=41%:%
%:%107=42%:%
%:%108=43%:%
%:%109=44%:%
%:%111=46%:%
%:%112=46%:%
%:%113=46%:%
%:%114=47%:%
%:%115=48%:%
%:%116=49%:%
%:%117=50%:%
%:%118=51%:%
%:%119=52%:%
%:%120=53%:%
%:%128=55%:%
%:%140=57%:%
%:%141=58%:%
%:%142=59%:%
%:%143=60%:%
